Automated troubleshooting toolset for SDNs

### 2013-10-27 draft.tex
NICE ended

For many years, the debugging have always been one of the major concerns in network maintenance. In order to localize problems into the network, operators usually use a narrow toolset, composed of traditional tools as \verb@ping@, \verb@traceroute@ and \verb@SNMP@ agents\cite{zeng2012automatic}.\\
Hopefully, this difficult and time-consuming\cite{zeng2012automatic} process could change, given the deployment of Software-Defined Networks. In a nutshell, SDNs are based on the separation between control- and data planes, which offer the opportunity of programmable networks\cite{mckeown2008openflow}. Those SDNs are composed of a network of switches managed by a logically-centralized controller, whose role is to (un-)install rules into the flow table of the switches, to read traffic statistics and respond to the network activity.\\
However, and because SDNs allows different operators and developers to dynamically program the same network, the complexity of software will increase\cite{Sherwood:2010:PNT:1924943.1924969} and potentially the numbers of bugs.
To minimise the trade-off between introducing new functionality and  increase the number of bugs in the network, there is a serious need for a complete and effective automated testing toolset, allowing the admins to focus on fixing the issues instead of localizing them. In traditional network architecture, it is almost impossible to create such an automated test suite, due to the complexity of \textit{knowing the operator's intent} and \textit{checking network behavior against intent}\cite{Heller:2013:LSL:2491185.2491197} (for more information about how traditional networks could be extended to support automated troubleshooting, please refer to \cite{Heller:2013:LSL:2491185.2491197}).\\
As we will explain (see Section~\ref{layering}), we use the different layers of the SDN stack to review the available tools for automated troubleshooting. Note that the methodology and the structure of this paper is influenced from the one used in\cite{Heller:2013:LSL:2491185.2491197}. It indeed seems the better way to articulate and present the different tools destined to localize the problems and their cause in an automated way.
The paper is structured as follow : first we recall the different layers of the SDN stack, and how it can be leveraged to provide automated troubleshooting for SDNs, then we present different existing tools by positioning them regarding the SDN layering, and we eventually conclude.

@@

\begin{figure}
\centering
\includegraphics[scale=0.65]{layers.png}
\label{layers}
\caption{SDN architecture\protect\footnotemark}
\end{figure}
\footnotetext{Source : \cite{Heller:2013:LSL:2491185.2491197}}
\section{Network Troubleshooting - Tools}
\label{Network Troubleshooting - Tools}
\subsection{Network Debugger (NDB)}

@@

Following the SDN layering technique, once the erroneous 2-layers correspondance has been determined, another tool can be used to spot the bug more precisely, either in the controller software or in the firmware running on the switches.\\
The NICE tool is used to produce traces leading to a bug, by systematically exploring the set of possible states of the system (including controller, switches and hosts) and checking them regarding to some network invariants\cite{Canini:2012:NWT:2228298.2228312}. What makes this contribution to OpenFlow applications verification a real asset is its ability to deal with a very large space state. The real challenge for testing openFlow applications is the scalability related to its wide environment : there is a potentially unbound state space, due to a large space of switch state, of inputs packets and of event orderings\cite{Canini:2012:NWT:2228298.2228312}. To handle that situation, developers can create abstractions of their application and use traditional model checking techniques to prove the properties about the system. Example tools are SPIN and JavaPathFinder, whose the mains limitations are respectively the complexity or writing the model due to the domain-specific information and a significant performance slackening\cite{Peresini:2011:YOA:2079327.2079345}. Moreover, both are suffering from some state-space explosion.\\

>>>>

For many years, the debugging has always been one of the major concerns in network maintenance. In order to localize problems into the network, operators usually use a narrow toolset, composed of traditional tools as \verb@ping@, \verb@traceroute@ and \verb@SNMP@ agents\cite{zeng2012automatic}.\\
Hopefully, this difficult and time-consuming\cite{zeng2012automatic} process could change, given the deployment of Software-Defined Networks. In a nutshell, SDNs are based on the separation between control- and data planes, which offer the opportunity of programmable networks\cite{mckeown2008openflow}. Those SDNs are composed of a network of switches managed by a logically-centralized controller, whose role is to (un-)install rules into the flow table of the switches, to read traffic statistics and respond to the network activity.\\
However, and because SDNs allows different operators and developers to dynamically program the same network, the complexity of software will increase\cite{Sherwood:2010:PNT:1924943.1924969} and potentially the numbers of bugs.
To minimise the trade-off between introducing new functionality and  increase the number of bugs in the network, there is a serious need for a complete and effective automated testing toolset, allowing the admins to focus on fixing the issues instead of localizing them. In traditional network architecture, it is almost impossible to create such an automated test suite, due to the complexity of \textit{knowing the operator's intent} and \textit{checking network behavior against intent}\cite{Heller:2013:LSL:2491185.2491197} (for more information about how traditional networks could be extended to support automated troubleshooting, please refer to \cite{Heller:2013:LSL:2491185.2491197}).\\
As we will explain (see Section~\ref{layering}), we use the different layers of the SDN stack to review the available tools for automated troubleshooting. Note that the methodology and the structure of this paper is influenced from the one used in\cite{Heller:2013:LSL:2491185.2491197}. It indeed seems the better way to articulate and present the different tools destined to localize the problems and their cause in an automated way.
The paper is structured as follow : first we recall the different layers of the SDN stack, and how it can be leveraged to provide automated troubleshooting for SDNs, then we present different existing tools by positioning them regarding the SDN layering, and we eventually conclude.

@@

\begin{figure}
\centering
\includegraphics[scale=0.60]{layers.png}
\label{layers}
\caption{SDN architecture\protect\footnotemark}
\end{figure}
\footnotetext{Source : \cite{Heller:2013:LSL:2491185.2491197}}
\section{Network Troubleshooting - Tools}
\label{Network Troubleshooting - Tools}
\subsection{Network Debugger (NDB)}

@@

Following the SDN layering technique, once the erroneous 2-layers correspondance has been determined, another tool can be used to spot the bug more precisely, either in the controller software or in the firmware running on the switches.\\
The NICE tool is used to produce traces leading to a bug, by systematically exploring the set of possible states of the system (including controller, switches and hosts) and checking them regarding to some network invariants\cite{Canini:2012:NWT:2228298.2228312}. What makes this contribution to OpenFlow applications verification a real asset is its ability to deal with a very large space state. The real challenge for testing openFlow applications is the scalability related to its wide environment : there is a potentially unbound state space, due to a large space of switch state, of inputs packets and of event orderings\cite{Canini:2012:NWT:2228298.2228312}. To handle that situation, developers can create abstractions of their application and use traditional model checking techniques to prove the properties about the system. Example tools are SPIN and JavaPathFinder, whose the mains limitations are respectively the complexity or writing the model due to the domain-specific information and a significant performance slackening\cite{Peresini:2011:YOA:2079327.2079345}. Moreover, both are suffering from some state-space explosion.\\
NICE extends traditional model checking with symbolic execution of the event-handler. It identifies therefore equivalence classes of packets that cover all code paths of the application. NICE combines also specific search strategies allowing it to reduce the state space by up to 20 times\cite{Canini:2012:NWT:2228298.2228312}.
